---
title: "HW10"
author: "Sarah Stover"
date: "2025-05-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
**Advanced ggplotting:** 
(Worked on this Homework with Rei)
Using the Palm Trees data set from TidyTuesday:

```{r}
#loading the packages
library(ggplot2)
library(ggridges)
library(ggbeeswarm)
library(GGally)
library(ggpie)
library(ggmosaic)
library(scatterpie)
library(waffle)
library(DescTools)
library(treemap)
library(devtools)
library(extrafont)

#reading in the Palm Tree data set
palmtrees <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-03-18/palmtrees.csv',show_col_types = FALSE)

#looking at top rows of df
head(palmtrees)
```

**Bar Plot**

looking at the species per subfamilies and the different fruit shapes per subfamily
```{r}
#barplot of subfamily species with different fruit shapes 
palm_barplot <- ggplot(palmtrees) + 
                          aes(x=palm_subfamily,fill=fruit_shape) +
                          geom_bar(position="dodge",color="black",size=0.5) +
                          labs(x = "Subfamily", 
                               y = " # of Species within Subfamily", 
                               fill = "Fruit Shape") +
                          scale_fill_brewer(palette = "Pastel1") +
                          theme(text = element_text(family = "Helvetica"),
                                  axis.title.x = element_text(size = 12),
                                  axis.text.x = element_text(size = 10),
                                  axis.title.y = element_text(size = 10))
                          
palm_barplot
ggsave(palm_barplot,file="palm_barplot.png",width=10,height=6,bg = "white")

```

**Pie Chart**

looking at the size of each subfamily of palm tree

```{r}
palm_piechart <- ggpie::ggpie(data=palmtrees,
                             group_key="palm_subfamily",
                             count_type="full",
                             label_info="ratio", 
                             label_type="none") +
                            scale_fill_brewer(palette = "Pastel1") +
                            labs(fill = "Subfamily Name") +
                            theme(text = element_text(family = "Helvetica"))
        
palm_piechart
ggsave(palm_piechart,file="palm_piechart.png",width=10,height=6,bg = "white")

```
**sub-setting the data set by sub-family**

```{r}
arecoideae <- palmtrees[which(palmtrees$palm_subfamily == "Arecoideae"),]
calamoideae <- palmtrees[which(palmtrees$palm_subfamily == "Calamoideae"),]
ceroxyloideae <- palmtrees[which(palmtrees$palm_subfamily == "Ceroxyloideae"),]
coryphoideae <- palmtrees[which(palmtrees$palm_subfamily == "Coryphoideae"),]
nypoideae <- palmtrees[which(palmtrees$palm_subfamily == "Nypoideae"),]
```
**Beeswarm Plot**
This beeswarm plot explores Petiole size and if the palm trees in the sub-family 
Calamoideae climb. Petiole is the stalk that connects the stem to the leaf blade. 
This shows that Calamoideae plants with a longer petiole tend to be climbers.
```{r}
palm_beeswarm <- ggplot(data=calamoideae) +
                            aes(x=climbing,y=max__petiole_length_m,color=climbing) + 
                            ggbeeswarm::geom_beeswarm(method = "center",size=3) +
                            scale_color_brewer(palette = "Accent") +
                            labs(x = "Climbing Type", y = "Maximum Petiole Length") +
                            theme(legend.position="none") +
                            theme(text = element_text(family = "Helvetica"),
                                  axis.title.x = element_text(size = 12),
                                  axis.text.x = element_text(size = 10),
                                  axis.title.y = element_text(size = 12))

palm_beeswarm
ggsave(palm_beeswarm,file="palm_beeswarm.png",width=10,height=6,bg = "white")

```

**Waffle plot**

This plot shows the number of tribes in the subfamily Calamoideae and the number of species per
tribe 

```{r}
tabled_data <- as.data.frame(table(class=coryphoideae$palm_tribe))

palm_waffle <- ggplot(data=tabled_data) +
                          aes(fill = class, values = Freq) +
                          waffle::geom_waffle(n_rows = 10, size = 0.3, colour = "white") +
                          scale_fill_manual(name = "Tribe Names:",
                                            values = c("#698B69", "#528B8B","#CD8C95", "#B2DFEE", "#EE9572","#CDCDC1","#CD96CD","#CDBA96"),
                                            labels = c("Borasseae", "Caryoteae", "Chuniophoeniceae","Corypheae","Cryosophileae","Phoeniceae","Sabaleae","Trachycarpeae")) +
                          coord_equal() +
                          theme_void() +
                          theme(legend.position = "bottom", text = element_text(family = "Helvetica"))

palm_waffle
ggsave(palm_waffle,file="palm_waffle.png",width=10,height=6,bg = "white")

```
**Ridgeline plot**

This plot shows the range of maximum stem heights per palm sub-family

```{r}
palm_subfamily_df <- data.frame(palmtrees$palm_subfamily)
palm_stemheight_df <- data.frame(palmtrees$max_stem_height_m)
palm_df <- cbind(palm_subfamily_df, palm_stemheight_df)
palm_df_clean <- palm_df[complete.cases(palm_df),]

palm_ridgeline <- ggplot(data=palm_df_clean,
                         aes(x=palmtrees.max_stem_height_m,y=palmtrees.palm_subfamily, fill = palmtrees.palm_subfamily)) + 
                         scale_fill_brewer(palette = "Accent") +
                         ggridges::geom_density_ridges(scale = 0.9) +
                         ggridges::theme_ridges() +
                         scale_x_continuous(limits = c(-10,100)) +
                         xlab("Maximum stem height (m)") +
                         ylab("Subfamily") +
                         theme(legend.position = "none", text = element_text(family = "Helvetica"))

palm_ridgeline
ggsave(palm_ridgeline,file="palm_ridgeline.png",width=10,height=6,bg = "white")

```






