---
title: "HW5"
author: "Sarah Stover"
date: "2025-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Question 1:

```{r}
n_dims <- runif(1, 3, 10) 
#vector with values from 1-n_dims^2
vec <- c(1:(n_dims^2))
#random sampling of vector above 
samp <- sample(vec)
#a square matrix with all the random samples that is n_dims length  
m <- matrix(data=samp, nrow=n_dims, ncol=n_dims)
print(m)
#transposing the matrix 
m_trans <- t(m)
print(m_trans)
```
The transposed matrix has the columns and rows switched from before. 
```{r}

#sum and mean first row
sum(m_trans[1,])
mean(m_trans[1,])

#sum and mean last row
sum(m_trans[n_dims,])
mean(m_trans[n_dims,])

#eigen function
eigen <- eigen(m_trans)
print(eigen)
typeof(eigen)
```
To my knowledge, the values that are returned when the eigen function is used tell you the factor used to scale the matrix or the relationship between the coefficients and the vectors. I think it returns a list of non-zero decomposition values hence "typeof" returning "list".

Question 2: 
```{r}
#
my_matrix <- matrix(data=runif(16), nrow=4, ncol=4)
my_logical <- c(runif(100)>50)
my_letters <- sample(letters, length(letters))
#putting all of these in 1 list 
my_list <- list(my_matrix, my_logical, my_letters)

# making a list of the second object in my matrix, logical, letters
new_list <- list(my_matrix[2,2], my_logical[2], my_letters[2])
#returns type of object for every object in new_list
sapply(new_list, typeof)

#concatenated list of all values in new_list
b <- c(my_list)
typeof(b)

```

Question 3

```{r}
# 26 random numbers 
my_unis <- runif(26, 0, 10)
#sampling of capital letters
my_letters <- sample(LETTERS, length(LETTERS)) 

#creating the data frame 
df <- data.frame(my_unis, my_letters)
#taking random rows and making them NAs
df[sample(nrow(df), 4), "my_unis"] <- NA
#which rows have NAs
which(is.na(df["my_unis"]))

#ordering the data set alphabetically by 2nd var
df[order(df$my_letters),]

#calc column mean for 1st var without the NAs
mean(df$my_unis, na.rm = T) 

```



