---
title: "HW6"
author: "Sarah Stover"
date: "2025-02-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Data Set Information and Summary:** 

For this homework, we are working with survey data of beekeepers in Vermont who are members VBA (Vermont Beekeepers Association). In this dataset they rated or commented their satisfation with certain areas of the organization. We want to understand the demographics of VBA and how the organization's resources are allotted. Who is needs more support and how can we be more inclusive to minorities or new beekeepers. Instead of generating our own data, we just built a nested forloop to go through the dataset and look at different categories. 

```{r}
#uploading and reading the data frame, getting
df <- read.csv("Annual Survey Results - 2025 Winter meeting.csv")
#exploring the df
head(df)

```

**Looking at Gendered differences in satisfaction between all organization categories:** 
```{r}
library(dplyr)
library(tidyr)

proportion_results <- list()

# Loop through each gender (Male and Female)
for (gender in c("Male", "Female")) {
  # Subset the data for this gender group
  gender_data <- df[df$Gender == gender, ]
  
  # Initialize a list to store proportions for this gender
  gender_proportions <- list()
  
  # Loop through each relevant column (Annual_meeting and Mentorship_programs)
  for (column in c("Annual_meetings", "Mentorship_program", "Access_resources", "Educational_workshops", "Industry_policy_insights", "Networking_opportunities", "News_updates", "Marketing_social")) {
    # Calculate the proportion of each category for this column
    prop_very_satisfied <- mean(gender_data[[column]] == "Very satisfied") * 100
    prop_satisfied <- mean(gender_data[[column]] == "Satisfied") * 100
    prop_unsatisfied <- mean(gender_data[[column]] == "Unsatisfied") * 100
    prop_very_unsatisfied <- mean(gender_data[[column]] == "Very unsatisfied") * 100
    prop_neutral <- mean(gender_data[[column]] == "Neutral") * 100
    prop_noexperience <- mean(gender_data[[column]] == "No prior experience") * 100
    
    # Store the results in the list for this column
    gender_proportions[[column]] <- c("Very satisfied" = prop_very_satisfied,
                                      "Satisfied" = prop_satisfied,
                                      "Unsatisfied" = prop_unsatisfied,
                                      "Very unsatisfied" = prop_very_unsatisfied,
                                      "Neutral" = prop_neutral,
                                      "No prior expreience" = prop_noexperience)
  }
  
  # Store the results for this gender
  proportion_results[[gender]] <- gender_proportions
}

# Print the results
print(proportion_results)

# Create a new column that indicates whether a person is unsatisfied or not
df$Mentorship_programClean <- ifelse(df$Mentorship_program %in% c("Unsatisfied", "Very Unsatisfied", "Neutral"),  "Unsatisfied", "Satisfied")

# Create a contingency table for gender vs unsatisfaction status
contingency_table <- table(df$Gender, df$Mentorship_programClean)

# Perform the Chi-squared test
chi_squared_test <- chisq.test(contingency_table)

# Print the results of the test
print(chi_squared_test)

print(contingency_table)
```
Because we are working with categorical values, we used a chi square analysis instead of ANOVA. Based on the p value (0.042) we can say there is a gendered difference between satisfaction levels. 

**Who is not coming to networking events by gender?:** 
```{r}
#library(ggplot2 

#removes NAs, mutates original dataset, creates a new column called networking_attendance. Then looks at df$Networking_opportunities to see if the answer is "no prior experience". if this is TRUE then its "did not attend" in the new column. Anything else is "Attended".
library(ggplot2)

df <- df %>%
  #filtering out NAs in Netwoking and Gender columns 
  filter(!is.na(Networking_opportunities) & !is.na(Gender) & Gender != "" & Networking_attendance != "") %>%
  mutate(Networking_attendance = ifelse(Networking_opportunities == "No prior experience", "Did Not Attend", "Attended"))

ggplot(df, aes(x = Gender, fill = Networking_attendance)) + 
  geom_bar(position = "dodge") + #side by side bars
  labs(title = "Networking Opportunity Attendence by Gender", x = "Gender", y = "count") + scale_fill_manual(values = c("Attended"= "pink", "Did Not Attend" = "orange"))
  
NetworkOpp_by_Gender_table <- table(df$Gender, df$Networking_attendance)
print(NetworkOpp_by_Gender_table)

```
There doesn't seem to be a bias in who is attending networking events by gender. The next thing to explore is by years of experience. Are those with less experience attending more or less than those already established?


**Participation in Networking Events by Beekeeping Experience:**
```{r}
# Create a new column that indicates whether a person is unsatisfied or not
df$Networking_opportunitiesClean <- ifelse(df$Networking_opportunities %in% c("Unsatisfied", "Very Unsatisfied", "Neutral"),  "Unsatisfied", "Satisfied")

# Create a contingency table for gender vs unsatisfaction status
contingency_table2 <- table(df$Beekeeping_experience, df$Networking_opportunitiesClean)

print(contingency_table2)

# Perform the Chi-squared test
chi_squared_test2 <- chisq.test(contingency_table2)


# Print the results of the test
print(chi_squared_test2)

# Get the standardized residuals
standardized_residuals2 <- chi_squared_test2$stdres

#Priint residuals
print(standardized_residuals2)
```
Based on the Chi square value, there no significant satisfaction based on years of experience. 
















