---
title: "HW8"
author: "Sarah Stover"
date: "2025-03-19"
output: html_document
---
Partner: Sydney Miller
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
**Simulating and Fitting Data Distributions**

Loading the data and libraries: 
  The data set we are using measures Honeybee hygenic behavior scores (a continuous dataset) 
```{r}
library(dplyr)

z <- read.csv("UBO_DataSheet.csv",header=TRUE,sep=",")
z <- z %>% 
  select(field_ID, assay_score) %>% 
  filter(!is.na(assay_score))

str(z)
summary(z)

colnames(z) <- c("ID", "myVar")

library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
# quick and dirty, a truncated normal distribution to work on the solution set

#z <- rnorm(n=3000,mean=0.2)
#z <- data.frame(1:3000,z)
#names(z) <- list("ID","myVar")
#z <- z[z$myVar>0,]
#str(z)
#summary(z$myVar)
```

Plot histogram of the data 
```{r}
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
#adds an empirical density curve
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
#get maximum likelihood parameters for normal distribution
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute
```

Plot normal probability density, can use stat function to generate probability density  
```{r}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
 p1 + stat
```
 
Plot exponential probability density
```{r} 
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
 p1 + stat + stat2
```
 
Plot uniform density probability 
```{r}
 stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
 p1 + stat + stat2 + stat3
```

Plot gamma probability density 
```{r}
 gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
```

Plot beta probability density 
```{r}
 pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial

```
**Which Probability Curve is the best fit?**
 
