---
title: "Strategic Coding Workshop"
author: "Sarah Stover"
date: "2025-03-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Outline: 
-problem of scaling 
-set upin Rstudio
-orgnization or Rstudio project 
-annotation of script files and data files
-logging 
-saving and restoring R data objects 
-useful tools (progress bar, timer, padded labels)
-user defined functions 
-pseudocodeand functional programming 


1. The problem of scaling 
symptoms- 
  copying and pasting many lines of code within a script 
  inconsistent formatting and obscure object/file names 
  little to no annotation 
  copying, renaming, editing entire script files
  copying # by hand from screen output 
  saving graphics through GUI 
  difficulty using code to do anything other than original task 
  eventually inability to reroduce original analysis 
  
What do we want?--> organized folders 
                    Annotation 
                    user defined functions  (short,simple,limited input and output)
                    handle multiple input/output files
                    use of RDS objects for complex and collaborative coding 
                    outputs saved as .csv or final format images 
                    relatively easy to add functionality
              
              
**Strategic Coding tools**              
```{r}

##load libraries --- 
library(pracma)
library(pryr)
library(devtools)
#this is how you install a package from github!
install_github("ngotelli/github")
#this will take you to the page where package functions are defined
help(package="upscalar")

## Subheader A ---

## Subheader B ---

library(package)
add_folder() 
add_folder(c("specialfolder1","specialfolders"))
metadata_template(file= "originalData/MyData.csv")
data_table_template(data_frame=NULL, file_name= "Outputs")

........

x <- runif(10) #an object to save
saveRDS(object = x,
        file="DataObjects/x.rds")#save to disk 
restored_x <- readRDS(file="DataObjects/x.rds") #reopen to new name 

```

